---
title: "SNSF Logistics Regression"
author: "Chiara Gilardi, Leslie O’Bray, Carla Schärer  and Tommaso Portaluri"
date: "6 April 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Load the functions and final data sets

```{r Load Functions and Data }
load("~/MSc Statistics/StatsLab/Analysis/SNFS Data/snsf_data.RData")
source("Cleaning Functions.R")
#source("Regressions.R")


# To access one single dataset:
    final.apps <- test$final.apps
    final.external <- test$final.reviews
    final.referees <- test$final.referees
    final.internal <- test$final.referees

library(caTools)
library(ROCR)
library(pROC)

```

# Exploratory Analysis

## Applications

## External Reviews

## Internal Reviews

# Regressions

### Regression with External Review Data set

Fit logistic regression using external criteria: goal is relevant importance of each criteria in this step.


```{r Logistic Regression External}

  # Calculate % female reviewers
    external_reviews_gender <- calculate_percent_female(final.external,
                                                        "ReviewerGender")
  
  # Select applications data we want to use
    external_regression_data <- final.apps[,c("IsApproved", "Age", "Gender", 
                                            "Division", "ProjectID")]
  
  # add in grades & Interaction
    average_ratings <- calculate_average_reviewers(final.external)
  
  # Merge applications data with external % females & average reviews
    external_regression_data <- merge(external_reviews_gender, 
                                    external_regression_data, by="ProjectID")
    external_regression_data <- merge(average_ratings, external_regression_data,
                                      by="ProjectID")
  
  # changing variables to factors
    # external_regression_data$ApplicantTrack<-as.factor(external_regression_data$ApplicantTrack)
    # external_regression_data$ScientificRelevance<-as.factor(external_regression_data$ScientificRelevance)
    # external_regression_data$Suitability<-as.factor(external_regression_data$Suitability)
    # external_regression_data$OverallGrade<-as.factor(external_regression_data$OverallGrade)


  # Function to fit a logistic Regression. Need to specify the data and Division
    # Options for divitions: "Div 1", "Div 2", "Div 3" and "All".
    # Split Ratio (default =0.8) If == 1 Then no Data will be split
    # cutoff (default 0.5)
    ExternalReviewModel<- function(data=external_regression_data,Div="All",
                                 SplitRatio=0.8, cutoff=0.5 ){ 
  
        if (Div == "All"){ 
          final.data<- data
        }
        else {
          final.data<- subset(data,Division==Div, select = -(Division)) 
        }
   
   
    #spliting the data into train and test set
      
      if (SplitRatio<1){
        split<-sample.split(final.data$IsApproved, SplitRatio = SplitRatio)
        Train<-subset(final.data, split=="TRUE")
        Test <-subset(final.data, split=="FALSE") 
      } else {
        Test<-Train<-final.data
      }
   
    # fitting the model
  
    #ExternalModel<-function(Train,Test,cutoff){
    
      # Cutoff
        cutoff<-cutoff
    
      # Optimize the model
        
        # full<- glm(IsApproved ~ .-(ProjectID), data=Train, family="binomial")
        # Model  <-step(full, direction = "backward", trace=0)
        # Model <- step(empty,scope=list(lower=empty,upper=full), direction="both",
        #     trace=0)
        Model <- glm(Train$IsApproved ~ .-(ProjectID),data=Train, 
                     family="binomial")
        
    
     # Testing the Treshold
        par(mfrow=c(1,2))
        predictor<-predict(Model, Test, type="response")
        ROCRPred<- prediction(predictor,Test$IsApproved)
        ROCRPerf<- performance(ROCRPred,"tpr","fpr")
        plot(ROCRPerf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1))
  
      # with respect to accuracy
        ROCRACC<- performance(ROCRPred,"acc")
        plot(ROCRACC)  
  
     # Confusion Matrices
        AccTable<-table(ActualValue=Test$IsApproved,Prediction=predictor>=cutoff)
        accuracy<-(sum(diag(AccTable))/sum(AccTable))
      
      # Return
        print(paste("Regresion for External Reviews.   ", "Division: ", Div))
        return(list(Model= summary(Model), 
                    `Confidence Intervals`=confint(Model),
                    `Confusion Matrix`=AccTable,
                    `Percentage of data used for Training`=paste(SplitRatio*100,"%"),
                    Accuracy=paste(round(accuracy,2)*100,"%")))
  
  
   }
  
```

## External Review

### All Divisions together

```{r All divisions}
ExternalReviewModel(data=external_regression_data, Div="All",cutoff=0.5,
                    SplitRatio=1) 
```


### Division 1

```{r External Div1}
Ext_Model<-ExternalReviewModel(data=external_regression_data, Div="Div 1",
                    SplitRatio=1,cutoff=0.5)
Ext_Model
```
### Division 2


```{r External Div2}
ExternalReviewModel(external_regression_data, Div="Div 2",
                    SplitRatio=1,cutoff=0.5)

```

### Division 3


```{r External Div3}
ExternalReviewModel(external_regression_data, Div="Div 3", 
                    SplitRatio=1, cutoff=0.5)
```

How probabilities change if gender is different

```{r}
beta_hat<-coefficients(Ext_Model$Model)[,1]
beta_hat

# probability
pfemale<- exp(beta_hat["Genderm"])/(1+exp(beta_hat["Genderm"])) # 56.8%
pmale<- exp(beta_hat["Genderm"]*2)/(1+exp(beta_hat["Genderm"]*2)) #63.4%
```

Division is also an influence.
```{r}
AppTrack<-1:6
pAppTrack<-exp(beta_hat["ApplicantTrack"]*AppTrack)/(1+exp(beta_hat["ApplicantTrack"]*AppTrack))
plot(AppTrack,pAppTrack, type="l")

```



### Regression with Internal Review Data set

Fit logistic regression using internal criteria: goal is relevant importance of each criteria in this step.


```{r Logistic Regression Internal}

  # Calculate % female reviewers
    internal_reviews_gender <- calculate_percent_female(final.internal, "RefereeGender")
  
  # Select applications columns we want to use
    internal_regression_data <- final.apps[,c("IsApproved", "ProjectID", "Gender",
                                            "Division","Age")]
  
  # add in grades & Interaction
    average_internal_ratings <- calculate_average_referee(final.internal)
  
  # Merge applications data with external % females & average reviews
    internal_regression_data <- merge(internal_regression_data, internal_reviews_gender, 
                                      by = "ProjectID")
    internal_regression_data <- merge(internal_regression_data, average_internal_ratings,
                                      by = "ProjectID")

  
  # changing variables to factors:
  internal_regression_data$Ranking <- factor(internal_regression_data$Ranking)
  internal_regression_data$ProjectAssessment <- factor(internal_regression_data$ProjectAssessment)
  internal_regression_data$ApplicantTrack <- factor(internal_regression_data$ApplicantTrack)


  # Function to fit a logistic Regression. Need to specify the data and Division
    # Options for divitions: "Div 1", "Div 2", "Div 3" and "All".
    # Split Ratio (default =0.8) If == 1 Then no Data will be split
    # cutoff (default 0.5)
    InternalReviewModel<- function(data=internal_regression_data,Div="All",
                                 SplitRatio=0.8, cutoff=0.5 ){ 
  
        if (Div == "All"){ 
          final.data<- data
        }
        else {
          final.data<- subset(data,Division==Div, select = -(Division)) 
        }
   
   
    #spliting the data into train and test set
      
      if (SplitRatio<1){
        split<-sample.split(final.data$IsApproved, SplitRatio = SplitRatio)
        Train<-subset(final.data, split=="TRUE")
        Test <-subset(final.data, split=="FALSE") 
      } else {
        Test<-Train<-final.data
      }
   
    # fitting the model
  
    #ExternalModel<-function(Train,Test,cutoff){
    
      # Cutoff
        cutoff<-cutoff
    
      # Optimize the model
        
        # full<- glm(IsApproved ~ .-(ProjectID), data=Train, family="binomial")
        # Model  <-step(full, direction = "backward", trace=0)
        # Model <- step(empty,scope=list(lower=empty,upper=full), direction="both",
        #     trace=0)
        Model <- glm(Train$IsApproved ~ .-(ProjectID),data=Train, 
                     family="binomial")
        
    
     # Testing the Treshold
        par(mfrow=c(1,2))
        predictor<-predict(Model, Test, type="response")
        ROCRPred<- prediction(predictor,Test$IsApproved)
        ROCRPerf<- performance(ROCRPred,"tpr","fpr")
        plot(ROCRPerf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1))
  
      # with respect to accuracy
        ROCRACC<- performance(ROCRPred,"acc")
        plot(ROCRACC)  
  
     # Confusion Matrices
        AccTable<-table(ActualValue=Test$IsApproved,Prediction=predictor>=cutoff)
        accuracy<-(sum(diag(AccTable))/sum(AccTable))
      
      # Return
        print(paste("Regresion for External Reviews.   ", "Division: ", Div))
        return(list(Model= summary(Model), 
                    #`Confidence Intervals`=confint(Model),
                    `Confusion Matrix`=AccTable,
                    `Percentage of data used for Training`=paste(SplitRatio*100,"%"),
                    `Accuracy`=paste(round(accuracy,2)*100,"%")))
 
  
   }
  
```


## Internal Review

### All Divisions together

```{r All divisions}
Int.all <- InternalReviewModel(data=internal_regression_data, Div="All",cutoff=0.5,
                    SplitRatio=1) 
Int.all
```


### Division 1

```{r Internal Div1}
Int_Model1<-InternalReviewModel(data=internal_regression_data, Div="Div 1",
                    SplitRatio=1,cutoff=0.5)
Int_Model1
```
### Division 2


```{r External Div2}
Int_Model2 <- InternalReviewModel(internal_regression_data, Div="Div 2",
                    SplitRatio=1,cutoff=0.5)
Int_Model2

```

### Division 3


```{r External Div3}
Int_Model3 <- InternalReviewModel(internal_regression_data, Div="Div 3", 
                    SplitRatio=1, cutoff=0.5)
Int_Model3
```

* All divisions: the odds of the project being approved decrease by 5.79% when the applicant is a male, assuming all the other variables to be constant.
* Division 1: the odds of the project being approved increase by 1.11% when the applicant is a male, assuming all the other variables to be constant.
* Division 2: the odds of the project being approved increase by 50.34% when the applicant is a male, assuming all the other variables to be constant.
* Divison 3: the odds of the project being approved decrease by 24.12% when the applicant is a male, assuming all the other variables to be constant.

(I got this results using grades as numeric, I'll change them later if needed)

```{r}
# All
beta_hat_all <- coefficients(Int.all$Model)["Genderm",1]
1-exp(beta_hat_all)

# Division 1
beta_hat_div1 <- coefficients(Int_Model1$Model)["Genderm",1]
exp(beta_hat_div1)

# Division 2
beta_hat_div2 <- coefficients(Int_Model2$Model)["Genderm",1]
exp(beta_hat_div2)

# Division 3
beta_hat_div3 <- coefficients(Int_Model3$Model)["Genderm",1]
1-exp(beta_hat_div3)
```


### Regression with Internal Review Data set

Fit logistic regression using cross-step criteria: goal is relevant importance of each step.


```{r Logistic Regression Board}


prepare_data_board_log_regression <- function(apps, internal, external) {
  
  # Extract columns from applications data
  board_regression_data <- final.apps[,c("IsApproved", "ProjectID", "Gender", "Division", "Age", "AmountRequested")]
  
  # Calculate average ratings for internal and external reviews
  average_internal_ratings <- calculate_average_referee(internal)[,c(1,4)]
  average_ratings <- calculate_average_reviewers(external)[,c(1,5)]
    
  # Merge with external reviews & referee data
  board_regression_data <- merge(board_regression_data, average_internal_ratings, by = "ProjectID")
  board_regression_data <- merge(board_regression_data, average_ratings, by = "ProjectID")
  board_regression_data$Ranking <- as.factor(board_regression_data$Ranking)
  board_regression_data$OverallGrade <- as.factor(board_regression_data$OverallGrade)
  
  # Select Division
  board_div<- subset(board_regression_data,Division=="Div 1", select = -Division)
  
}

board_data <- prepare_data_board_log_regression(final.apps, internal = final.internal, external = final.external)

board_log_regression <- glm(board_data$IsApproved ~ .-(ProjectID), family="binomial", data = board_data)

BoardModel<- function(data=board_data,Div="All",
                                 SplitRatio=0.8, cutoff=0.5 ){ 
  
        if (Div == "All"){ 
          final.data <- data
        }
        else {
          final.data<- subset(data,Division==Div, select = -(Division)) 
        }
   
   
    #spliting the data into train and test set
      
      if (SplitRatio<1){
        split<-sample.split(final.data$IsApproved, SplitRatio = SplitRatio)
        Train<-subset(final.data, split=="TRUE")
        Test <-subset(final.data, split=="FALSE") 
      } else {
        Test<-Train<-final.data
      }
   
    # fitting the model
  
    #ExternalModel<-function(Train,Test,cutoff){
    
      # Cutoff
        cutoff <- cutoff
    
      # Optimize the model
        
        # full<- glm(IsApproved ~ .-(ProjectID), data=Train, family="binomial")
        # Model  <-step(full, direction = "backward", trace=0)
        # Model <- step(empty,scope=list(lower=empty,upper=full), direction="both",
        #     trace=0)
        Model <- glm(Train$IsApproved ~ .-(ProjectID),data=Train, 
                     family="binomial")
        
    
     # Testing the Treshold
        par(mfrow=c(1,2))
        predictor<-predict(Model, Test, type="response")
        ROCRPred<- prediction(predictor,Test$IsApproved)
        ROCRPerf<- performance(ROCRPred,"tpr","fpr")
        plot(ROCRPerf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1))
  
      # with respect to accuracy
        ROCRACC<- performance(ROCRPred,"acc")
        plot(ROCRACC)  
  
     # Confusion Matrices
        AccTable<-table(ActualValue=Test$IsApproved,Prediction=predictor>=cutoff)
        accuracy<-(sum(diag(AccTable))/sum(AccTable))
      
      # Return
        print(paste("Regresion for External Reviews.   ", "Division: ", Div))
        return(list(Model= summary(Model), 
                    #`Confidence Intervals`=confint(Model),
                    `Confusion Matrix`=AccTable,
                    `Percentage of data used for Training`=paste(SplitRatio*100,"%"),
                    `Accuracy`=paste(round(accuracy,2)*100,"%")))
 
  
   }
 
  
board.all <- BoardModel(data=board_data, Div="All",
                    SplitRatio=1,cutoff=0.5)

```


```{r Janine External Regression}

head(final.external)

# Calculate % female reviewers
    external_reviews_gender <- calculate_percent_female(final.external,
                                                        "ReviewerGender")
  
  # Select applications data we want to use
    external_regression_data <- final.apps[,c("IsApproved", "Age", "Gender", 
                                            "Division", "ProjectID")]
  
  # add in grades & Interaction
    average_ratings <- calculate_average_reviewers(final.external)
  
  # Merge applications data with external % females & average reviews
    external_regression_data <- merge(external_reviews_gender, 
                                    external_regression_data, by="ProjectID")
    external_regression_data <- merge(average_ratings, external_regression_data,
                                      by="ProjectID")
    
    # combine suitability & project proposal
    external_regression_data$ProposalCombined <- round((external_regression_data$ScientificRelevance+
                                                            external_regression_data$Suitability)/2, 0)
    

    # Simple first regression: Overall Grade 4-6 >- "Good", 1-3 -> bad
    external_regression_data$SimplifiedOverallGrade <- ifelse(external_regression_data$OverallGrade < 4, 0, 1)
    external_regression_data$SimplifiedProposal <- ifelse(external_regression_data$ProposalCombined < 4, 0, 1)
    external_regression_data$SimplifiedApplicant <- ifelse(external_regression_data$ApplicantTrack < 4, 0, 1)

    
    # Janine regression 
        
    janine_reg_proposal <- glm(SimplifiedProposal ~ Age + Gender + Division + 
                                  PercentFemale + Gender:PercentFemale, data=external_regression_data, family="binomial")
    
    summary(janine_reg_proposal)

    janine_reg_applicant <- glm(SimplifiedApplicant ~ Age + Gender + Division + 
                                  PercentFemale + Gender:PercentFemale, data=external_regression_data,  family="binomial")
  


        
    # Leslie's probably wrong interpretation of Janine's idea
    leslie_reg_proposal <- glm(SimplifiedOverallGrade ~ Age + Gender + Division + 
                                 ProposalCombined + PercentFemale, data=external_regression_data, family="binomial")
    
    leslie_reg_applicant <- glm(SimplifiedOverallGrade ~ Age + Gender + Division + 
                                 ApplicantTrack + PercentFemale, data=external_regression_data, family="binomial")
      
    

    

```

```{r Janine Internal Regression}

# Calculate % female reviewers
    internal_reviews_gender <- calculate_percent_female(final.internal,
                                                        "RefereeGender")
  
  # Select applications data we want to use
    internal_regression_data <- final.apps[,c("IsApproved", "Age", "Gender", 
                                            "Division", "ProjectID")]
  
  # add in grades & Interaction
    average_ratings <- calculate_average_referee(final.internal)
  
  # Merge applications data with external % females & average reviews
    internal_regression_data <- merge(internal_reviews_gender, 
                                    internal_regression_data, by="ProjectID")
    internal_regression_data <- merge(average_ratings, internal_regression_data,
                                      by="ProjectID")
    

    # Simple first regression: Overall Grade 4-6 >- "Good", 1-3 -> bad
    internal_regression_data$SimplifiedApplicantTrack <- ifelse(internal_regression_data$ApplicantTrack < 4, 0, 1)
    internal_regression_data$SimplifiedProjectAssessment <- ifelse(internal_regression_data$ProjectAssessment < 4, 0, 1)


    
    # Janine regression 
        
    janine_reg_project <- glm(SimplifiedProjectAssessment ~ Age + Gender + Division + 
                                  PercentFemale + Gender:PercentFemale + Gender:Division, data=internal_regression_data)
    
    summary(janine_reg_project)
    confint(janine_reg_project)

    janine_reg_track <- glm(SimplifiedApplicantTrack ~ Age + Gender + Division +
                              Gender:Division+PercentFemale + Gender:PercentFemale, data=internal_regression_data)
    
    summary(janine_reg_track)
    confint(janine_reg_track)

```